---
title: "ICC World Cup Case Study"
author: "Bhavya Gupta"
date: "Monday, September 25, 2017"
output: word_document
---

### Problem statement

Use the past ICC World Cup data available at cricinfo.

Set up hypotheses to disprove the following Null hypotheses:

* Tendulkar's  batting average is not better than Rahul Dravid's.
* The Strike rates of Tendulkar and Dravid are the same
* Percentage of of time Tendulkar gets out in single digits is the same as Dravid's
* Dravid is more consistent than Tendulkar

Finally, set up a problem that might allow you to do paired t test. 


``` {r}
df_Rahul_Dravid <- 
  data.frame(Series = c("ICC World Cup (in England/Ireland/Netherlands/Scotland), 1999",
                        "ICC World Cup (in Kenya/South Africa/Zimbabwe), 2002/03",
                        "ICC World Cup (in West Indies), 2006/07"), 
             Innings = c(8,10,3), 
             avg = c(65.85, 63.6, 40.5),
             SR = c(85.52, 64.11, 72.32))

df_SR_Tendulkar <- 
  data.frame(Series = 
    c("Benson & Hedges World Cup (in Australia/New Zealand), 1991/92",
      "Wills World Cup (in India/Pakistan/Sri Lanka), 1995/96",
      "ICC World Cup (in England/Ireland/Netherlands/Scotland), 1999",
      "ICC World Cup (in Kenya/South Africa/Zimbabwe), 2002/03",
      "ICC World Cup (in West Indies), 2006/07",
      "ICC Cricket World Cup (in Bangladesh/India/Sri Lanka), 2010/11"), 
  Innings = c(7, 7, 7, 11, 3, 9), 
  avg = c(47.16, 87.16, 42.16, 61.18, 32, 53.55),
  SR = c(84.73, 85.87, 90.03, 89.25, 110.34, 91.98))

#avg_df <- data.frame(SR = df_SR_Tendulkar$avg[c(3,4,5)], RD =df_Rahul_Dravid$avg )

```


#### Load required libraries

```{r}

library(ggplot2)

## install scales package prior to use

library(scales)

library(car)

library(gplots)

```

#### Write Useful functions

``` {r}
test_equalvar <- function(sample1, sample2) {
  val <- c(sample1,sample2)
  grp <- as.factor(c(rep(1,length(sample1)), rep(2,length(sample2))))
  out <- levene.test(val,grp)
  return(out["Pr(>F)"][1]) 
}

perform_t <- function(sample1,sample2) {
  cat("\n\n")
  print("Null Hypothesis: Samples Variances are equal")
  cat("\n")
  print("Alternative hypothesis: Sample variances are not equal")         
  cat("\n\n")
  p <-   test_equalvar(sample1, sample2)[1,1]
  print(paste(" P value calculated by Levene\'s test is: ",p))
#
  if (p > 0.05)  {
    print(paste("We conclude the variances are equal - Null hypothesis"))         
    res <- t.test(sample1,sample2,alternative="less",var.equal = TRUE)
  } else {
    print(paste("We conclude the variances are not equal - Alternative hypothesis"))         
    res <- t.test(sample1,sample2,alternative="less",var.equal = FALSE)
  }   
  
  return(res)
      
}    

```

##### 1) Batting average 

* Tendulkar's  batting average is not better than Rahul Dravid's.

 1 Test equality of variances


In **statistics**, *Levene's test* is an inferential statistic used to assess the equality of variances for a variable calculated for two or more groups. It tests the null hypothesis that the population variances are equal (called homogeneity of variance or homoscedasticity). If the resulting p-value of Levene's test is less than some significance level (typically 0.05), the obtained difference in sample variances are unlikely to have occurred based on random sampling from a population with equal variances.

* Null Hypothesis: Tendulkar's  batting average is not better than Rahul Dravid's

* Alternative Hypothesis: Tendulkar's  batting average is better than Rahul Dravid's

``` {r}

sample1 <- df_SR_Tendulkar$avg
sample2 <- df_Rahul_Dravid$avg
p_val <- perform_t(sample1,sample2)$p.value  
decision <- ifelse (p_val > 0.05,paste("Since p value : ",p_val, " > 5% Significance level -- > We accept Null hypothesis"), paste("Since p value : ",p_val, "< 5% significance level --> We reject Null hypothesis"))
cat("\n\n Observation is ",decision)
# Visual display

par(mfrow = c(1,2))
boxplot(df_SR_Tendulkar$avg, main = "Batting Average of Tendulkar")
boxplot(df_Rahul_Dravid$avg, main = "Batting Average of Rahul Dravid")

###
```

##### Observations

Since p-value = 0.0.41566 at we accept the null hypothesis at confidence level 95% and conclude **Tendulkar's  batting average is not better than Rahul Dravid's**

##### 2) Strike Rate 

* The Strike rates of Tendulkar and Dravid are the same

* Null Hypothesis: -The Strike rates of Tendulkar and Dravid are the same

* Alternative Hypothesis: -The Strike rates of Tendulkar and Dravid are not same

``` {r}

sample1 <- df_SR_Tendulkar$SR
sample2 <- df_Rahul_Dravid$SR
p_val <- perform_t(sample1,sample2)$p.value  
decision <- ifelse (p_val > 0.05,paste("Since p value : ",p_val, " > 5% Significance level -- > We accept Null hypothesis"), paste("Since p value : ",p_val, "< 5% significance level --> We reject Null hypothesis"))
cat("\n\n Observation is ",decision)

# Visual display

par(mfrow = c(1,2))
boxplot(df_SR_Tendulkar$SR, main = "Strike Rate of Tendulkar")
boxplot(df_Rahul_Dravid$SR, main = "Strike Rate of Rahul Dravid")
###
```

##### Observations

Since p-value = .9824 at we accept the null hypothesis at confidence level 95% and conclude **The Strike rates of Tendulkar and Dravid are the same**

##### 3) Single Digit getting out

* Percentage of of time Tendulkar gets out in single digits is the same as Dravid's

* Null Hypothesis: -The Strike rates of Tendulkar and Dravid are the same

* Alternative Hypothesis: -Percentage of of time Tendulkar gets out in single digits is the same as Dravid's

``` {r}
single_digit <-matrix(c(2,19,9,35),ncol=2)
colnames(single_digit) <- c('Rahul Dravid','Tendulkar')


rownames(single_digit) <- c('Less than 10','Other')

result.prop <- prop.test(single_digit)

cat("\n P-value: ", result.prop$p.value)

```

##### Observations

Since p-value = 0.456 at we accept the null hypothesis at confidence level 95% and conclude **Percentage of of time Tendulkar gets out in single digits is the same as Dravid's**


##### 4) Batting consistency

A coefficient of variation (CV) is a statistical measure of the dispersion of data points in a data series around the mean. The coefficient of variation represents the ratio of the standard deviation to the mean and it is a useful statistic for comparing the degree of variation from one data series to another, even if the means are drastically different from one another. 

It is calculates as follows:

CV = Standard Deviation / Mean value

http://www.investopedia.com/terms/c/coefficientofvariation.asp

We are going to test the following null hypothesis:

* Dravid is more consistent than Tendulkar

* Null Hypothesis: -Dravid is more consistent than Tendulkar

* Alternative Hypothesis: -Dravid is not more consistent than Tendulkar

``` {r}

Sachin_runs <- c(35,11,0,81,4,84,14,0,70,90,137,3,31,65,28,0,2,22,0,45,16,  
		 52,36,81,152,50,98,5,97,15,83,4,7,0,0,28,120,38,27,111,2,53,85,18)	

Dravid_runs <- c(54,13,0,145,53,2,61,29,17,1,0,62,0,32,0,0,0,47,14,0,60)
##
SD_Sachin_runs   <- sd(Sachin_runs)
SD_Dravid_runs   <- sd(Dravid_runs)
Mean_Sachin_runs <- mean(Sachin_runs)
Mean_Dravid_runs <- mean(Dravid_runs)
COV_Sachin	 <- SD_Sachin_runs / Mean_Sachin_runs
COV_Dravid	 <- SD_Sachin_runs / Mean_Dravid_runs

cat("\n Co-variance of Sachin: ", COV_Sachin,"\n",
    "Co-variance of Dravid: ", COV_Dravid,"\n")

conclusion <- ifelse(COV_Sachin > COV_Dravid, "Dravid is more consistent because of lower covariance","Sachin is more consistent because of lower covariance")

print(conclusion)
```

##### Observations

**Sachin is more consistent than Tendulkar because of lower co-variance**


##### 5)  Finally, set up a problem that might allow you to do paired t test. 

We collect data of number of fours scored by both Sachin and Dravid during the ODIs they played together in ICC World cup.

The paired sample t test is a statistical procedure used to determine whether the mean difference between two sets of observations is zero.


* Null Hypothesis: -Dravid and Tendulkar has hit equal number of fours 

* Alternative Hypothesis: -Number of Fours hit by Dravid and Tendulkar are not equal 

##### Arrange data

``` {r}

Fours <-matrix(
  c('ODI # 1444',5,5,
    'ODI # 1457',  10, 16,
    'ODI # 1463',17,0,
    'ODI # 1467',6,2,
    'ODI # 1473',0,0,
    'ODI # 1476',4,5,
    'ODI # 1480',5,2,
    'ODI # 1948',0,7,
    'ODI # 1951',0,3,
    'ODI # 1957',2,10,
    'ODI # 1969',3,8,
    'ODI # 1975',2,12,
    'ODI # 1983',3,1,
    'ODI # 1985',2,7,
    'ODI # 1988',7,3,
    'ODI # 1992',0,5,
    'ODI # 1993',2,1,
    'ODI # 2538',0,1,
    'ODI # 2542',0,2,
    'ODI # 2550',6,0),byrow = T, nrow=20)
Fours
colnames(Fours) <- c("Match","Dravid","Sachin")

```


#####  Analysis of Boundary scored


``` {r}

n = 20
df <- as.data.frame(Fours)
df$Dravid <- as.numeric(df$Dravid)
df$Sachin <- as.numeric(df$Sachin)

data_means <- c(mean(df$Dravid), mean(df$Sachin))
data_SD    <- c(sd(df$Dravid),sd(df$Sachin))
data_SE    <- c(sd(df$Dravid),sd(df$Sachin)) / sqrt(n)

barplot2(data_means, 
    ci.l = data_means - data_SE, 
    ci.u = data_means + data_SE, 
    plot.ci = TRUE, ci.lty = 1, ci.color = 1, ci.lwd = 2, ci.width = 0.5, 
    col = c("grey", "green"), lwd = 2, width = 0.25, 
    xlab = "Batsmen", ylab = "Average Fours", 
    names.arg = c("Sachin", "Dravid"), 
    axis.lty = 1, main = "Error Bars =  +/- SE")

  p <-   test_equalvar(df$Dravid, df$Sachin)[1,1]
  print(paste(" P value calculated by Levene\'s test is: ",p))
#
  
```


###### Observation

Variances are equal.

### Conduct a paired t test

``` {r}

  if (p > 0.05)  {
    print(paste("We conclude the variances are equal"))         
    p_val <- t.test(df$Dravid,df$Sachin,paired=TRUE, alternative = "two.sided",var.equal = T)$p.value
  } else {
    print(paste("We conclude the variances are not equal"))         
    p_val <- t.test(df$Dravid,df$Sachin,paired=TRUE, alternative = "two.sided",var.equal = FALSE)$p.value

  }   


cat("\n p-value: ", p_val,"\n")

compare <- ifelse(p_val > 0.05, "We accept null hypothesis","We reject null hypothesis")

print(compare)

```

###### Observation

Since the p-value of 0.2026127 is > 5% (level of significance), we accept the null hypothesis at 95% level of confidence and conclude that
**Dravid and Tendulkar has hit equal number of fours**.
